---
title: "Aides Drug"
author: "Estrella Aguilar "
date: "2/13/24"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

```{r}
drugies <- filter(triUmp,DrugUser == "yes")
```



# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drugies)
barchartGC(~infected + treatment,data=drugies, type="percent")

```

Almost 80% of the people who reported to have aids have taken a placebo not the real drug. That shows that the drug seems to be working in preventing aids 



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment,data=drugies)
rowPerc(table1)
colPerc(table1)
```

There were three times the people who were in the placebo group who reported getting aids in comparison to the people who were taking the real drug.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

we ran three inferential test in this section, two chi- squared tests, and a fishers exact test. In the Chi squared test we are getting a p value of 0.0000001242 which is far lower than 0.05 & a  chi-squared equal 27.955 (approx.28) which is far greater than our critical value for degrees of freedom in case 1. Given these scores we can infer that there is a difference in the data 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We can see through our graphical and numerical results. The placebo group contracted the disease at a greater frequency than the treatment group. This coupled with the tests of statistical inference show taht there is a statistically significant difference in the data. We conclude that there is a negative correlation between the drug use, and the disease frequency. Due to the high number of people included in the study, we can assume that the random variation latest small role, if any in the data, we can infer from the the results, assuming the data was fairly correct that there is accusation in the case and the drug use is effective.


